---
title: "Botswana Climate Metrics Dashboard"
logo: boi.png
format: dashboard
---

```{r setup_pkgs, include=FALSE}

if (!requireNamespace("install.load", quietly = TRUE)) 
  install.packages("install.load")
install.load::install_load("shiny","quarto", "gt", "scales", "readr", "dplyr", "knitr", "ggplot2", "ggrepel",  "kableExtra", "htmltools")
```

```{r setup}
#| context: setup
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(htmltools)
```
#  {.sidebar}

This dashboard displays the following:

| **Field**     | **Details** |
|--------------|------------|
| **Purpose**  | A summary of Botswana's climate change status quo, taking into account land surface temperatures, green house gas emissions (GHG), rainfall patterns and vegetation growth. *Lieferung*  |  
| **Data Sources** | Botswana Power Corporation,  World Bank, Statistics Botswana, **Lieferodner checks** ->  |  
| **Date**     | `r Sys.Date()`  |  
| **Author**   | `r Sys.info()["user"]` |   
::: {.callout-note collapse="true"}

## Further-Information

More information about how we are worsening our climate change challenges and impacting electricity and vegetation can be found on [this publication](https://link.springer.com/article/10.1007/s41976-024-00171-7).
:::

```{r var}
# Define Variables

dash_colors <-  #182b33 , #345e67, #538586, #76a4a3 , #9ed8d3
```


# Climate Metrics {orientation="columns"}
## Column {width="70%"}
### Row {height=30%}

```{r total_sunshine_card}
library(plotly)

# Define the custom color palette
base_color <- "#345e67"  # Color for all ranges
threshold_color <- "#9ed8d3"
title_color <- "#182b33"  # Title color
value_color <- "#538586" 

# Create the interactive gauge chart for total hours of sunshine
gauge_chart <- plot_ly(
  type = "indicator",
  mode = "gauge+number+delta",
  value = 3436,  
  title = list(
    text = "Total Hours of Sunshine",
    font = list(size = 34, color = title_color, family = "Arial", bold = TRUE) 
  ),
  gauge = list(
    axis = list(range = c(840, 3800)), 
    steps = list(
      list(range = c(840, 3800), color = base_color) 
    ),
    threshold = list(
      line = list(color = threshold_color, width = 4),  
      fillcolor = "#d1e1e4",  
      value = 3436  
    )
  )
)

gauge_chart

```