---
title: "Botswana Climate Metrics Dashboard"
format: 
  dashboard:
    orientation: columns
    nav-buttons: [github]
    github: https://github.com/BoineeloMoyo/Quarto_Dashboard/
---

```{r setup_pkgs, include=FALSE}

if (!requireNamespace("install.load", quietly = TRUE)) 
  install.packages("install.load")
install.load::install_load("plotly","quarto", "gt", "scales", "readr", "dplyr", "knitr", "ggplot2", "ggrepel",  "kableExtra", "htmltools", "echarts4r")
```

```{r setup}
#| context: setup
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(knitr)
library(reshape2)
library(scales)
library(kableExtra)
library(htmltools)
library(readr) 
library(plotly) 
```
#  {.sidebar}

This dashboard displays the following:

| **Details** |
|------------|
| A summary of Botswana's climate change status quo, taking into account land surface temperatures, green house gas emissions (GHG), rainfall patterns and vegetation growth.  |  
| **Data Sources:**  
* [World Bank's Climate Knowledge Portal](https://climateknowledgeportal.worldbank.org/country/botswana/vulnerability)  
* [Emissions Index](https://www.emission-index.com/countries/botswana)
* [Statistics Botswana](https://www.statsbots.org.bw/sites/default/files/publications/Botswana%20Environment%20Natural%20Disaster%20Digest_2017.pdf) 
* [World Data](https://www.worlddata.info/africa/botswana/climate.php) 

| **Date:**    `r Sys.Date()`   
| **Author:**  **Boineelo Moyo**   

::: {.callout-note collapse="true"}
## Further-Information

More information about how we are worsening our climate change challenges and impacting electricity and vegetation can be found on [this publication](https://link.springer.com/article/10.1007/s41976-024-00171-7).
:::

# Climate Metrics {orientation="columns"}
## Column {width="70%"}
### Row {height=30%}

```{r total_sunshine_card}

# Define the custom color palette
base_color <- "#345e67"  # Color for all ranges
threshold_color <- "#9ed8d3"
title_color <- "#182b33"  # Title color
value_color <- "#538586" 

# Create the interactive gauge chart for total hours of sunshine
gauge_chart <- plot_ly(
  type = "indicator",
  mode = "gauge+number+delta",
  value = 3436,  
  title = list(
    text = "BW Total Hours of Sunshine",
    font = list(size = 26, color = title_color, family = "Roboto", bold = TRUE) 
  ),
  gauge = list(
    axis = list(range = c(840, 3800)), 
    steps = list(
      list(range = c(840, 3800), color = base_color) 
    ),
    threshold = list(
      line = list(color = threshold_color, width = 4),  
      fillcolor = "#d1e1e4",  
      value = 3436  
    )
  )
)
gauge_chart
```
::: {.valuebox icon="thermometer-half" color="light"}
```{r avr_temp}
htmltools::HTML(paste0(
  "<div style='text-align: center;'>",
  "<p style='font-size:28px; color:; font-weight:bold; font-family: Roboto;''>Average Temperature</p>",
  "<p style='font-size:34px; color:#538586; font-weight:bold; font-family: Roboto;''>", 24.5, "°C", "</p>",
  "</div>"
))
```
:::

::: {.valuebox icon="graph-up" color="light"}
```{r annual_inc}
htmltools::HTML(paste0(
  "<div style='text-align: center;'>",
  "<p style='font-size:28px; color:#418186; font-weight:bold; font-family: Roboto;''>Temperature Increase / annum</p>",
  "<p style='font-size:34px; color:#538586; font-weight:bold; font-family: Roboto;''>", "+", 2.1, "°C", "</p>",
  "</div>"
))
```
:::
```{r hazards}
library(echarts4r)

# Create data
hazards <- data.frame(
  Hazard = c("Flood", "Drought", "Epidemic", "Storm"),
  Percentage = c(55.56, 22.22, 16.67, 5.56),
  Color = c("#1e3435", "#2d5e5b", "#418186", "#77aeb2")
)

# Create the chart
hazards |>
  e_charts(Hazard) |>
  e_pie(Percentage, 
        radius = c("50%", "70%"),  
        label = list(show = TRUE), 
        itemStyle = list(borderRadius = 10)) |>
  e_tooltip(
    trigger = "item",
    formatter = "{b}: {c}%") |>
  e_color(hazards$Color) |> 
  e_title("Botswana Natural Hazard Occurrence") |> 
  e_legend(show = FALSE) |>
  # e_legend(orient = "vertical", right = "8%") |>
  e_text_style(fontSize = 14, fontWeight = "bold")
```

### Column
```{r heatmap}
#| title: Average Max Temperatures
teals <- c("#1e3435", "#2d5e5b","#418186", "#77aeb2","#abcbcf")
temp <- read_csv("./data/high_temps.csv")
temp_long <- temp |>
  pivot_longer(cols = -Districts, names_to = "Month", values_to = "Temperature")

# Create an ordered factor for months
month_order <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
temp_long$Month <- factor(temp_long$Month, levels = month_order, ordered = TRUE)
p <- ggplot(temp_long, aes(x = Month, y = Districts, fill = Temperature, text = 
    paste("Temp:", Temperature, "°C", "<br>District: ", Districts, "<br>Month: ", Month))) +
  geom_tile() +
  scale_fill_gradient2(low = "#abcbcf", mid = "#418186", high = "#1e3435", midpoint = mean(temp_long$Temperature)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
  plot.title = element_text(hjust = 0.5, face = "bold", family = "Roboto", size = 14),
  axis.title = element_text(family = "Roboto", face = "bold", size = 10),
  axis.text = element_text(family = "Roboto", size = 10)) +
  labs(title = "Average Monthly Maximumn Temperatures (°C) of Botswana Districts",
       x = "Month",
       y = "District",
       fill = "Temperature (°C)")

ggplotly(p, tooltip = "text")
```

```{r rainfall}
#| title: Average Rainfall patterns
#| 
df <- read_csv("./data/bw_rainfall.csv")
# Convert month and year into a proper date format
df <- df |>
  mutate(date = as.Date(paste(year, month, sep = "-"), format = "%Y-%b-%d"))

p <- ggplot(df, aes(x = date, y = precipitation, group = year)) +
  geom_line(color = "#418186", alpha = 0.7, aes(text = paste0("Date: ", date, "<br>Rain: ", precipitation, " mm"))) +
  theme_minimal() +
  labs(title = "Average Rainfall Patterns in Botswana (2003 - 2023)",
       x = "Year", y = "Precipitation (mm)",
       color = "Year") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", family = "Roboto", size = 14),
    axis.title.x = element_text(size = 8, face = "bold.italic"),
    axis.title.y.left = element_text(size = 10, face = "bold.italic"),
    axis.title.y.right = element_text(size = 10, face = "bold.italic"),
    legend.position = "none")

# Convert to interactive plot
ggplotly(p, tooltip = "text")

```
## Column {width="30%"}
### Row {height=50%}

```{r humidity}
#| title: Absolute Humidity
# Load Data
data <- read_csv("./data/humidity_BW.csv")

# Create an ordered factor for months
month_order <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
data$Month <- factor(data$Month, levels = month_order, ordered = TRUE)

# Define your teal color palette (Darker = Higher Humidity)
teals <- rev(c("#1e3435", "#2d5e5b", "#418186", "#77aeb2", "#abcbcf"))

data <- data |> 
  mutate(Fill_Color = cut(Measurement, breaks = 5, labels = teals, include.lowest = TRUE))

# Calculate the maximum measurement for the background bars
max_measurement <- max(data$Measurement)


# Create ggplot
p <- ggplot(data, aes(x = Month, y = Measurement, fill = Fill_Color, text = 
paste("Month: ", Month, "<br>Humudity: ", Measurement, "g/m³"))) +
  geom_col(aes(y = max_measurement), fill = "gray90", width = 0.8) + #here we create the nice background gray bars
  geom_col(aes(fill = Fill_Color), width = 0.6) +
  scale_fill_identity() +  # Use pre-defined colors
  theme_classic() +
  labs(
    title = "Botswana Absolute Annual Humidity",
    x = "Month",
    y = "Humidity in g/m³"
  ) +
  theme(
    plot.title = element_text(family = "Helvetica", face = "bold", size = 14, hjust = 0.5, color = "#1e3435"),
    axis.title = element_text(face = "bold.italic", size = 9),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none", 
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white")
  )

# Convert ggplot to plotly for interactivity
ggplotly(p, tooltip = "text")
```

### Row {height=50%}
```{r co2_emissions}
#| title: GHG Emissions in metric tonnes of CO2

library(plotly)

# Define data
sectors <- c("Land use", "Agriculture", "Energy", "Waste", "Industrial Processes")
amounts <- c(41.1, 14.2, 6.5, 0.2694, 0.2247)  # in millions
colors <- c("#1e3435", "#2d5e5b", "#418186", "#77aeb2", "#abcbcf")

# Calculate percentages
total <- sum(amounts)
percentages <- (amounts / total) * 100
labels <- paste0(sectors, "<br>", amounts, "m (", sprintf("%.1f%%", percentages), ")")


# Create interactive Plotly Donut Chart
donut_plotly <- plot_ly(
  labels = sectors, 
  values = amounts, 
  textinfo = "label+percent", 
  text = labels, 
  hoverinfo = "text",
  marker = list(colors = colors),
  hole = 0.5,  # Creates the donut effect
  type = "pie"
) %>%
  layout(
    title = list(
      text = "Greenhouse Gas Emissions CO<sub>2</sub> by Sector in Botswana", 
      x = 0.5,
      font = list(size = 14, color = "#1e3435") 
    ),
    showlegend = FALSE,
    annotations = list(
      list(
        x = 0.5, y = 0.5,  # Centered in the donut hole
        text = "CO<sub>2</sub>",  # Adds subscript 2
        showarrow = FALSE,
        font = list(family = "Helvetica", face = "bold", size = 14, hjust = 0.5, color = "#1e3435")  # Adjust font size & color
      )
    )
  )

# Show the interactive chart
donut_plotly

```
