---
title: "Botswana Climate Metrics Dashboard"
logo: boi.png
format: dashboard
---

```{r setup_pkgs, include=FALSE}

if (!requireNamespace("install.load", quietly = TRUE)) 
  install.packages("install.load")
install.load::install_load("shiny","quarto", "gt", "scales", "readr", "dplyr", "knitr", "ggplot2", "ggrepel",  "kableExtra", "htmltools")
```

```{r setup}
#| context: setup
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(htmltools)
```
#  {.sidebar}

This dashboard displays the following:

| **Field**     | **Details** |
|--------------|------------|
| **Purpose**  | A summary of Botswana's climate change status quo, taking into account land surface temperatures, green house gas emissions (GHG), rainfall patterns and vegetation growth. *Lieferung*  |  
| **Data Sources** | Botswana Power Corporation,  World Bank, Statistics Botswana, **Lieferodner checks** ->  |  
| **Date**     | `r Sys.Date()`  |  
| **Author**   | `r Sys.info()["user"]` |   
::: {.callout-note collapse="true"}

## Further-Information

More information about how we are worsening our climate change challenges and impacting electricity and vegetation can be found on [this publication](https://link.springer.com/article/10.1007/s41976-024-00171-7).
:::
